Massively Parallel Monte Carlo Methods for Discrete Linear and Nonlinear Systems

### 2012-11-02 doc/prelim/presentation/prelim_presentation.tex
Removing stationary method slides
pull out some projection method slides

%%---------------------------------------------------------------------------%%
\begin{frame}{Stationary Methods}

  \begin{itemize}
  \item General stationary methods are formed by splitting the linear
    operator
  \end{itemize}

  \[
  \ve{A} = \ve{M} - \ve{N}
  \]

  \[
  \ve{x} = \ve{M}^{-1}\ve{N}\ve{x} + \ve{M}^{-1}\ve{b}
  \]

  \begin{itemize}
  \item We identify $\ve{H} =\ve{M}^{-1}\ve{N}$ as the
    \textit{iteration matrix}
  \end{itemize}

  \[
  \ve{x}^{k+1} = \ve{H}\ve{x}^{k} + \ve{c}
  \]

%%---------------------------------------------------------------------------%%
\begin{frame}{Stationary Methods Convergence}

  \begin{itemize}
  \item The qualities of the iteration matrix dictate convergence
  \item Define $\ve{e}^k = \ve{x}^k-\ve{x}$ as the error at the
    $k^{th}$ iterate
  \end{itemize}

  \[
  \ve{e}^{k+1} = \ve{H} \ve{e}^k
  \]

  \begin{itemize}
  \item We diagonalize $\ve{H}$ to extract its Eigenvalues
  \end{itemize}

  \[
  ||\ve{e}^{k}||_2 = \rho(\ve{H})^k ||\ve{e}^0||_2
  \]

  \begin{itemize}
  \item We bound $\ve{H}$ by $\rho(\ve{H}) < 1$ for convergence
  \end{itemize}

@@

    The residual of the system will always be \textit{minimized} with
    respect to the constraints
    \[
    ||\tilde{\ve{r}}||_2 \leq ||\ve{r}_0||_2,\ \forall \ve{r}_0 \in
    \mathbb{R}^N
    \]
  \end{beamerboxesrounded}

%%---------------------------------------------------------------------------%%
\begin{frame}{Putting It All Together}

  \begin{itemize}
  \item Choose $\ve{V}$ as a basis of $\mathcal{K}$ and $\ve{W}$ as a
    basis of $\mathcal{L}$
  \end{itemize}

  \[
  \boldsymbol{\delta} = \ve{V}\ve{y},\ \forall \ve{y} \in \mathbb{R}^N
  \]

  \[
  \ve{y} = (\ve{W}^T\ve{A}\ve{V})^{-1}\ve{W}^T\ve{r}_0
  \]

  \pause
  \begin{beamerboxesrounded}[upper=boxheadcolor,lower=boxbodycolor,shadow=true]
    {Projection Method Iteration}
    \[
    \ve{r}^k = \ve{b} - \ve{A}\ve{x}^k
    \]
    \[
    \ve{y}^k = (\ve{W}^T\ve{A}\ve{V})^{-1}\ve{W}^T\ve{r}^k
    \]
    \[
    \ve{x}^{k+1} = \ve{x}^k + \ve{V}\ve{y}^k
    \]
    \[
    Update\ \ve{V}\ and\ \ve{W}
    \]
  \end{beamerboxesrounded}

@@

  \[
  \mathcal{K}_m(\ve{A},\ve{r}_0) = span\{\ve{r}_0, \ve{A}\ve{r}_0,
  \ve{A}^2\ve{r}_0, \dots, \ve{A}^{m-1}\ve{r}_0\}
  \]
  \[
  \mathcal{L} = \ve{A} \mathcal{K}_m(\ve{A},\ve{r}_0)
  \]

  \begin{itemize}
  \item Yields the normal system $\ve{A}^T\ve{A}\ve{x} =
    \ve{A}^T\ve{b}$
  \item Must generate an orthonormal basis $\ve{V}_m \in \mathbb{R}^{N
    \times m}$ for $\mathcal{K}_m(\ve{A},\ve{r}_0)$
  \item $\ve{W}_m = \ve{A}\ve{V}_m$
  \item Typically choose a Gram-Schmidt-like procedure such as Arnoldi
    or Lanzcos
  \end{itemize}

%%---------------------------------------------------------------------------%%
\begin{frame}[fragile]{GMRES}

  \begin{algorithm}[H]
    \begin{algorithmic}[1]
      \STATE $\ve{r}_0 := \ve{b}-\ve{A}\ve{x}_0$ 
      \STATE $\beta := ||\ve{r}_0||_2$ 
      \STATE $\ve{v}_1 := \ve{r}_0 / \beta$
      \COMMENT{Create the orthonormal basis for the Krylov subspace}
      \FOR{$j = 1, 2, \cdots, m$} 
      \STATE $\ve{w}_j := \ve{A}\ve{v}_j$
      \FOR{$i = 1, 2, \cdots, j$} 
      \STATE $h_{ij} \leftarrow \langle \ve{w}_j,\ve{v}_j \rangle$ 
      \STATE $\ve{w}_j \leftarrow \ve{w}_j - h_{ij}\ve{v}_i$ 
      \ENDFOR 
      \STATE $h_{j+1,j} \leftarrow ||\ve{w}_j||_2$ 
      \STATE $\ve{v}_{j+1} \leftarrow \ve{w}_j / h_{j+1,j}$ 
      \ENDFOR 
      \COMMENT{Apply the orthogonality constraints}
      \STATE $\ve{y}_m \leftarrow argmin_y ||\beta \ve{e}_1 -
      \ve{H}_m\ \ve{y}||_2 $ 
      \STATE $\ve{x}_m \leftarrow \ve{x}_0 + \ve{V_m} \ve{y}_m$
    \end{algorithmic}
    \caption{GMRES Iteration}
  \end{algorithm}

>>>>

@@

    The residual of the system is \textit{minimized} with respect to
    the constraints
    \[
    ||\tilde{\ve{r}}||_2 \leq ||\ve{r}_0||_2,\ \forall \ve{r}_0 \in
    \mathbb{R}^N
    \]
  \end{beamerboxesrounded}

@@

  \[
  \mathcal{K}_m(\ve{A},\ve{r}_0) = span\{\ve{r}_0, \ve{A}\ve{r}_0,
  \ve{A}^2\ve{r}_0, \dots, \ve{A}^{m-1}\ve{r}_0\}
  \]

  \begin{itemize}
  \item For GMRES:
  \end{itemize}
  \[
  \mathcal{L} = \ve{A} \mathcal{K}_m(\ve{A},\ve{r}_0)
  \]

  \begin{itemize}
  \item Yields the normal system $\ve{A}^T\ve{A}\ve{x} =
    \ve{A}^T\ve{b}$
  \item Must generate an orthonormal basis $\ve{V}_m \in \mathbb{R}^{N
    \times m}$ for $\mathcal{K}_m(\ve{A},\ve{r}_0)$
  \item Typically choose a Gram-Schmidt-like procedure such as Arnoldi
    or Lanzcos
  \item Short and long recurrence relations available for
    orthogonalization
  \end{itemize}
